Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2018-08-31 Response.tex

        In this paper, the natural, uncontrolled dynamics represented by stable manifold are considered as a good starting point for low thrust orbital transfers, as orbital maneuvers with low thrust over a short time period would not differ from those substantially. 
        But, as pointed out by the reviewer, the proposed approach does not impose any optimality in the constructed maneuver. 
        The scope of this paper is constructing low thrust orbital maneuvers without relying on massive computations. 
        To clarify these, the revised manuscripts states that there is no optimality guaranteed.

    \item
        \begin{itshape}

            * The extra analysis behind Fig 8 is a step in the right direction, but in
            practice it is hard to make sense of, and might require additional
            figures/better color choices.  Why are all the reachable sets nearly
            straight lines, when we expected some kind of closed path to arise from
            sweeping 0 < theta < 360?  Why are some progressions of sets so irregular
            (e.g. red cross --> blue cross --> green cross --> yellow cross signifying
            increase in travel time with fixed large umax, which results in darting
            back and forth in the x dimension)?  The 25 lines of discussion do not
            acknowledge these oddities, nor do they seem to completely line up with the
            figure.  There seem to remain unexplored/unexplained relationships that
            would in practice need to be used to justify parameter choices by anyone
            applying this method.   
        \end{itshape}

        The various reachable sets appear as straight lines due to the discretization used and the level of zoom of the figure.
        In reality the reachable sets are highly elongated ellipses on the \Poincare section.
        This is due to a combination of factors, namely that the low thrust propulsion cannot greatly deviate from the natural dynamics and that the nature of the optimization results in approximate numerical solutions subject to the choice of tolerance levels.
        In addition, as the figure shows there is a nontrivial relationship between terminal time and control magnitude. 
        The resulting reachability set is heaviliy dependent on the specific problem, choice of initial conditions, terminal time, and control bound.
        The figure is meant to illustrate some large scale, general properties of the approach, namely that an incrase in terminal time and/or control magnitude will have the effect of enlarging and moving the resulting reachable set.
     The intention is not to present an exhaustive search/discussion of this specific situation but rather to illustrate the effect of variations in terminal time and maximum control magnitude.
        The revised manuscript explicitly states this:
        \begin{correction}
            The reachable sets presented in~\cref{fig:varying_tf_reachability_sets} are highly dependent on the initial condition, terminal time, and maximum control bound. 
            This example demonstrates that increasing the \( u_m \) results in a larger displacement between the controlled and uncontrolled trajectories.
            The reachable set is enlarged from a single point, as shown in~\cref{fig:manifold_poincare} by the black point, to a larger region as shown in~\cref{fig:varying_tf_poincare}.
            The choice of \( u_m\), \( t_f \) and initial condition all combine to change the resulting reachability set. 
        \end{correction}
    \item
        \begin{itshape}

            * As mentioned before, the Poincare illustrations of Fig 2 and 3 would be
            more helpful if they were more general.  The first could benefit from
            showing nonperiodic dynamics, the second from showing a non-circular
            reachable boundary (I know the text acknowledges this but please update the
            figure).  
        \end{itshape}
        The illustrations of the \Poincare section match the typical depication in many other publications~\cite{koon2001}.
        Furthermore, the illustrations are merely utilized as helpful visualizations of the concepts, specifically periodic solutions, lower dimensional \Poincare sections, and reachable sets, rather than specific depictions of a solution.
        An additional comment is added to the caption specifically mentioning that the reachable set can be any arbitrary shape, and potentially non-continuous.

@@

    \item
        \begin{itshape}

            * I wonder under what circumstances the reachable set might intersect a
            given $\theta_d$ in multiple places.  This is something that could be looked
            into to expand the strength of a submission focusing only on the
            single-horizon use of the method.
        \end{itshape}

        Thank you for the comment.
        This is an interesting question and a possible avenue for future research.

@@

    \item 
        \begin{itshape}

            * There are still some very long single paragraphs in the paper.
        \end{itshape}

        Thank you for the comment. 
        The manuscript has been revised and the long single paragraphs have been divided.
\end{itemize}

### 2018-08-31 manuscript.tex
Update manuscript

A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0.2 & 0\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
The step size of the variational integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior in comparison to first order integration techniques without a significant increase in computational demand.
The variational integrator is constructed according to a discrete version of Hamilton's principle, specifically for the given dynamic system, and it provides long-term structural stability in capturing the effects of low-thrust propulsion systems.
General purpose numerical integrators, such as adaptive stepsize Runge-Kutta methods may yield a more accurate state trajectory depending on the choice of integration parameters, but at the expense of additional computational demand and total energy deviations.
However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.

@@

        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
        % place
    \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Reachability set on a \Poincare section: Pictorial representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
            The terminal state, \( x_n\), is the intersection without any control input. 
            Adding a control input allows for the terminal state, \( x_f \), to be displaced by some distance/cost \( J \) as measured on the section.
            We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
            Computation of the maximum distance, or reachability, for a variety angles gives a discrete approximation of the reachability set.\label{fig:reachability_set}}
\end{figure}

>>>>

A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0.2 & 0\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
The step size of the variational integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.

\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.

This work uses a second order integrator which has improved energy behavior in comparison to first order integration techniques without a significant increase in computational demand.
The variational integrator is constructed according to a discrete version of Hamilton's principle, specifically for the given dynamic system, and it provides long-term structural stability in capturing the effects of low-thrust propulsion systems.
General purpose numerical integrators, such as adaptive stepsize Runge-Kutta methods may yield a more accurate state trajectory depending on the choice of integration parameters, but at the expense of additional computational demand and total energy deviations.
However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.

@@

        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
        % place
    \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Reachability set on a \Poincare section: Pictorial representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
            The terminal state, \( x_n\), is the intersection without any control input. 
            Adding a control input allows for the terminal state, \( x_f \), to be displaced by some distance/cost \( J \) as measured on the section.
            We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
            Computation of the maximum distance, or reachability, for a variety angles gives a discrete approximation of the reachability set.
        In general, the reachable set can be an arbitrary shape on the section rather than the circular set depicted.\label{fig:reachability_set}}
\end{figure}
